---
title: "Applied Problem Set 2"
author: "Sabrina Li"
date: "29/03/2020"
output: 
  pdf_document:
    
    number_sections: yes
  html_document:
    df_print: paged
urlcolor: blue
---
```{r message=FALSE, warning=FALSE}
library(tidyverse)
Sys.setlocale(locale = 'USA')
# install.packages('testthat')
library(testthat)
library(tidyverse)
library(lubridate)

```

<!-- .Rmd files use  markdown, a text mark up language, to provide formating.--> 
<!--Text include within these strange arrows are comments and will not show up when you knit-->

# Front matter
This submission is my work alone and complies with the 30535 integrity policy.

Add your initials to indicate your agreement: **__**

Add your collaborators: **__**

Late coins used this pset: 0. Late coins left: X. 
<!--You may use up to two for a given assignment.)-->

# Problems

## Part 1

### 1.1

####  1
```{r}
# Part 1

## 1
system.time({
  
  sp = read.csv('parking_tickets_one_percent.csv')
  
})



test_that('row number', expect_equal(nrow(sp), 287458))

```

####  2

```{r}

infs = file.info('parking_tickets_one_percent.csv')
size_mb = infs$size / 1024 /1024
size_mb # mb


size_mb * 100 # population size


```


####  3

it seems ordered by issue_date (ascending), but there is exceptations.

```{r}
sp %>% head()

(sp$issue_date %>% ymd_hms() %>% diff() < 0) %>% sum()
```


####  4

```{r}

sapply(sp, function(x) sum(is.na(x))) %>% sort() -> missbycol
missbycol
```

####  5

```{r}
missbycol %>% tail(3)

```

- zipcodes may be privacy

- existing of other two ones are not necessary

### 1.2

####  1
```{r}
library(lubridate)

sp$issue_date %>% ymd_hms() -> sp$ymd

(sp$ymd %>% year() == 2007) %>% sum() # 2007


table(sp$ymd %>% year()) # different years
```

About 200,000 tickets were issued in both 2011 and 2014, for example, and revenue increased from about $21 million to $25 million. There were also similar numbers of tickets issued in 2007 and 2015 — about 250,000 tickets. Revenue jumped from about $25 million to $32 million.

The difference is not so large, but there is descending trend.



####  2

```{r}
t20 = sp$violation_description %>%  table() %>% sort() %>% tail(20) %>% names()
sp %>%  filter(violation_description %in% t20) %>%
  ggplot(aes(violation_description, 
             fill = violation_description)) + 
  geom_bar() + theme(axis.title.x=element_blank(),
                     axis.text.x=element_blank(),
                     axis.ticks.x=element_blank()) + labs(title = 'bar plot')
                                  
```





### 1.3

####  1
```{r}
sp$unit %>% is.na() %>% sum() # nums of unit missing 
```

####  2

```{r}
uit = read.csv('unit_key.csv', skip = 2)

# uit  %>% head()
uit$Reporting.District %>% length()
uit$unit = uit$Reporting.District %>% as.numeric() # units
uit[uit$unit %>% is.na() == F,] -> uit
uit$unit %>% unique() %>% length() # nums of units

```


####  3

```{r}

left_join(sp, uit, by = 'unit') -> joinunit
sapply(uit, function(x) sum(is.na(x))) %>% 
  sort() # check missing value

joinunit$Department.Name %>% is.na() %>% 
  sum() # rows in the tickets data do not have a match

nrow(joinunit) - joinunit$Department.Name %>% 
  is.na() %>% sum() # # rows in the tickets data have a match

(uit$unit %in% joinunit$unit) %>%
  sum() # rows in the unit table have a match

(uit$unit %in% joinunit$unit == F) %>%
  sum()  # rows in the unit table do not  have a match



```

#### 4

```{r}
lapply(sp$unit_description, function(x){
  strsplit(x, '-')[[1]][1]
}) %>% unlist() -> joinunit$unit_gen


table(joinunit$unit_gen)[1:2] # count for two orgs.

joinunit[joinunit$unit_gen == 'CPD',] ->sub_chi
sub_chi$Department.Description %>% table() %>% sort() %>% 
  tail(5) # top five depts in Chicago

```




### 1.4


```{r}
zipdf = read.csv('chi_zips.csv')
colnames(zipdf) ='GEOID'
library(tidycensus)

census_api_key("f51cd09bd6d0f8cdc620d3c64a5f7893c2efebe1")
lkup = load_variables(year = 2017,dataset = "acs5")
lkup$concept %>% grepl( pattern = 'black', x = ., ignore.case = T) %>% which() -> ids
lkup[ids,] -> xx

acs_data <- get_acs(
  geography = "zip code tabulation area",
  variables = c("B01001_001","B19013_001", "B01001B_001"),
  year = 2019,
  geometry = FALSE
)  # down load

acs_data %>% head()

# acs_data$GEOID = as.integer(acs_data$GEOID)
colnames(acs_data)[1] = 'zipcode'

cens = pivot_wider(acs_data,id_cols = zipcode, 
                   names_from = variable, values_from = estimate)
colnames(cens)[2:4] =c('population', 'income', 'balck')
cens$balck = cens$balck / cens$population

cens %>% head() # clean 
cens_join = left_join(joinunit, cens, by = 'zipcode' )

cens_join %>% filter(ticket_queue != 'Paid') %>% filter(zipcode %>% is.na() == F)%>% 
  group_by(zipcode) %>% 
  summarise(count = n()) %>% arrange(desc(count)) %>% 
  head(3) %>% select(zipcode) # top 3 neigborhoods

```



## Part 2

#### 2.1.1

```{r}
sp[sp$ticket_queue == 'Notice',] -> notpay
double_pay = ((notpay$current_amount_due + 
                 notpay$total_payments) >=  2* notpay$fine_level1_amount) 
table(double_pay) # not hold for all

not_not = notpay[double_pay == F,]
not_100 =not_not[(not_not$current_amount_due +
                    not_not$total_payments) >= 100, ]
head(not_100) # sample

```

it does Not hold for all. if not unpaid, price will doubled, and have other fees.

#### 2.1.2

```{r, dev='png'}
notpay$notice_level %>% table() %>% sort() %>%
  rev() %>% names() -> proces
proces # process
# install.packages('DiagrammeR')
DiagrammeR::grViz("digraph {

graph[layout = dot, rankdir = LR]

SEIZ
DLS
FINL
VIOL
DETR

SEIZ -> DLS -> FINL ->VIOL ->DETR
}")

```

I assume the later the process, the fewer the notices numbers.


#### 2.1.3

```{r}
sp %>% group_by(violation_description) %>%
  summarise(n_code = unique(violation_code)%>%
              length()) %>% filter(n_code > 1) ->multicode

multicode

sp[sp$violation_description %in% multicode$violation_description ,] -> mutidf

table(mutidf$violation_description, mutidf$violation_code)

```




### 2.1.4

```{r}
sp %>% group_by(violation_code) %>% 
  summarise(n_desc = unique(violation_description)%>% 
              length()) %>% filter(n_desc > 1) ->multidesc

multidesc

sp[sp$violation_code %in% multidesc$violation_code ,] -> mutidf

table( mutidf$violation_code, mutidf$violation_description)
```



### 2.1.5

```{r}
table(sp$violation_description) %>%
  sort() %>% tail(50)
```

for OBSTRUCT ROADWAY is too general and there is lots of detailed violation,for example, PARK OR STAND IN VIADUCT/UNDERPASS 


### 2.2.1

```{r}
sp$violation_description %>% 
  grepl('NO CITY STICKER', . , ignore.case = T) -> ids
citystk = sp[ids, ]
citystk$violation_description %>% unique() # codes

citystk = citystk[citystk$violation_description != 
                    'NO CITY STICKER VEHICLE OVER 16,000 LBS.',]
table(citystk$ymd %>% year(), citystk$violation_code) -> time_table
time_table # old new codes according time.

citystk %>% group_by(violation_code) %>% 
  summarise(init_fee = unique(fine_level1_amount)) # initial price.


```



### 2.2.2

```{r}

time_table %>% apply(., 1, sum) -> time_table2
plot(names(time_table2), time_table2, type = 'l', 
     xlab = 'time', ylab = 'count',
     main = 'number of missing sticker tickets  over time')



```



### 2.2.3

```{r}
citystk %>% ggplot(aes(ymd, fine_level1_amount, 
                       color = violation_code)) + 
  geom_line() + labs(title = 'cost in time')



```

about from 2012 spring.


### 2.2.4

```{r}

citystk[citystk$ymd %>% year() == 2011,] %>% nrow() * (200 - 120)


```
revenue increases.

### 2.2.5

```{r}

citystk$yr = citystk$ymd %>% year()
citystk[citystk$yr %in% c(2011, 2012),] -> vorafter

vorafter$paid = vorafter$ticket_queue == 'Paid'

vorafter %>% group_by(yr, paid) %>% 
  summarise(count = n()) %>% 
  pivot_wider(id_cols = yr, names_from = paid, values_from = count) -> cont

cont$rate = cont[,3] / (cont[,2] + cont[,3])

cont$pay = cont$rate * c(120, 200)

cont

```



### 2.2.6

```{r}

vorafter = citystk

vorafter$paid = vorafter$ticket_queue == 'Paid'

vorafter %>% group_by(yr, paid) %>% 
  summarise(count = n()) %>% 
  pivot_wider(id_cols = yr, names_from = paid, values_from = count) -> cont

cont$rate = cont[,3] / (cont[,2] + cont[,3]) %>% unlist() %>% as.numeric()
cont$rate = cont$rate[,1]
cont = as.data.frame(cont)
colnames(cont) = c('year', 'pay', 'no', 'rate')
cont %>% ggplot(aes(year, rate)) + 
  geom_line() +labs(title = 'repayment rate') +geom_vline(xintercept = 2012, 
                                                          linetype="dotted", 
                color = "blue", size=1.5)


```


The rate decreases over time, the degree increase largely after introducing new price.

### 2.2.7

```{r}

sp = cens_join
sp[sp$ymd %>%year() == 2011, ] -> df11

df11$violation_code %>% table() %>% sort() %>% tail(3)
df11$violation_description %>% table() %>% sort() %>% tail(3)

df11$pay = df11$ticket_queue == 'Paid'
df11 %>% group_by(violation_code) %>% 
  summarise(recv = sum(pay) / n() * unique(fine_level1_amount) ) %>%
  arrange(recv) %>% tail(3) -> top3
top3

```



### 2.2.8

```{r}
df11$high_income = df11$income >  median(df11$income, na.rm = T)

df11 = df11[df11$zipcode %>% is.na() == F,]
df11 = df11[df11$high_income %>% is.na() == F,]
diff_df = df11 %>% group_by(high_income, violation_code) %>% 
  summarise(count = n()) %>% 
  pivot_wider(id_cols = violation_code, 
              names_from = high_income , values_from = count)

diff_df$diff = diff_df$`TRUE` - diff_df$`FALSE`
diff_df %>% arrange(diff %>% desc()) %>% head(3) -> rich_top3
colnames(rich_top3)[2:3] =c('low', 'high')
rich_top3 %>% pivot_longer(cols = c('low', 'high')) %>% 
  ggplot( aes(x = violation_code , y = value, fill = name)) +
  geom_bar(stat = "identity", position = 'dodge') +
  labs(title = 'most different violations')



citystk = df11[df11$violation_code %in% rich_top3$violation_code,]
citystk$yr = citystk$ymd %>% year()
# citystk[citystk$yr %in% c(2011, 2012),] -> vorafter
vorafter = citystk
vorafter$paid = vorafter$ticket_queue == 'Paid'

vorafter %>% group_by(yr, paid) %>% summarise(count = n()) %>% 
  pivot_wider(id_cols = yr, names_from = paid, values_from = count) -> cont
as.numeric(cont[,3] * 80 )


```
